---
title: "index"
author: "Steven Litvin"
date: "November 30, 2017"
output:
  html_document:
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Scatterplot
```{r cars}

library(tidyverse)
g <- ggplot(data = mpg, aes(x=displ, y=hwy))
g + geom_point(aes(shape = class))
g + geom_point(aes(color = class))
g + geom_point(aes(color = year, size = cty, alpha = 0.2))

```

## Geoms

```{r geom_expample}
ggplot(mpg, aes(x=displ, y=hwy))+
geom_point(aes(x=displ, y=hwy))+
geom_smooth(aes(x=displ, y=hwy))+
  labs(title="test")
```

## data wrangling with 'dplyr'
```{r}
library(tidyverse)
gapminder <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")
```

## useful functions to get to know your data:
```{r}
head(gapminder) #show first 6 rows of data
tail(gapminder) #show last 5 rows of data
head(gapminder, 30)
str(gapminder) #checks if variables are num, int, char ext
names(gapminder) # column headers
dim(gapminder) #dimentions of the data set
ncol(gapminder)#num of columns 
nrow(gapminder)#num of row
c(nrow(gapminder), ncol(gapminder)) # c is combine
summary (gapminder) #summary stats or each columns
```

## for everything above, we were operating on the whole dapminder dataset ... $ is for a column

head(gapminder$pop)

##actually working with dplyr
```{r}
filter(gapminder, country == "Mexico")
sweden <- filter(gapminder, country == "Sweden")
mean(sweden$lifeExp)
```

##pipe operator '%>%'
gapminder %>% head(3)

```{r}
gapminder %>% 
  filter(country == "Sweden") %>% 
  summarize (mean_lifeExp = mean(lifeExp))
```

#select 2 columns
```{r}
select(gapminder, year, lifeExp)
gapminder %>% 
  select(year, lifeExp)
```
#lets filter some more
```{r}
gapminder %>% 
  filter(country == "Cambodia") %>% 
  select (-continent, -lifeExp)
```

#lets filter EVEN more
```{r}
gapminder %>% 
  filter(country == "Mexico",
         year == 2002)
```


#mutate () to add columns
```{r}
gapminder %>%
  mutate(gdp= pop*gdpPercap)
```

#lets add an index
```{r}
gapminder %>%
  mutate(index = 1:nrow(gapminder)) %>%
  tail()
```

#lets add an index and dpp and save it all to a new array
```{r}
expanded_data <- gapminder %>%
  mutate(index = 1:nrow(gapminder), gdp= pop*gdpPercap)
```

#find maximum gdp of Egpyt and Vietnam - Create new column, but its one value which will be used to populate every data line
```{r}
gapminder %>%
  filter(country %in% c("Egypt", "Vietnam")) %>%
  mutate(max_gdpPercap = max(gdpPercap))
```

#ind maximum gdp of Egpyt and Vietnam seperately- Create new column, report the max for each country
```{r}
gapminder %>%
  filter(country %in% c("Egypt", "Vietnam")) %>%
  group_by(country) %>%
  summarize(max_gdpPercap = max(gdpPercap))
```

#ind maximum gdp of every countryseperately- Create new column, report the max for each country
```{r}
gapminder %>%
  group_by(country) %>%
  summarize(max_gdpPercap = max(gdpPercap))
```

#combining data sets 
```{r}
## read in CO2 emmisionts data
library(tidyverse)
carbon <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/co2.csv")

carbon %>% head()

##create new variablie with only gapminder 2007 data
gapminder <- readr::read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")
gap_2007 <- gapminder %>%
  filter(year==2007)
gap_2007 %>% head()
gap_2007 %>% str()

#left_join gap_2007 and carbon
lj <- left_join(gap_2007, carbon, by = "country")
lj %>% dim()

##righ_join same datasets
rj <- right_join(gap_2007, carbon, by = "country")
rj %>% dim()
```